Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-02 paper.tex
edits

Imagine an agent faced with a suite of new problems totally different
from anything it has seen before. It has at its disposal a basic set
of primitive actions it can compose to build solutions to these problems, but
it is no idea what kinds of primitives are appropriate for which
problems nor does it know the higher-level vocabulary in
which solutions are best expressed.
How can our agent get off the ground?

The AI and machine learning literature contains two broad takes on this problem.
The first take is that the agent should come up with a better representation of the space of solutions,
for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
Our contribution is a general algorithm for fusing these two takes on the problem:
we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
alongside a bottom-up discriminative model that regresses from problems to solutions.
We evaluate our algorithm on four domains:
building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
We show that \system can construct a set of basis primitives suitable for discovering solutions in each of these domains

@@

Our key observation is that the generative and recognition models can bootstrap off of each other,
greatly increasing the tractability of BPL.

@@

    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
  \end{scope}
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

The goal of \system is to both induce a DSL and find good programs solving each of the tasks.
Our strategy is to iterate through three steps: (1) searching for programs that
solve the tasks, (2) learning a better neural recognition model -- which we use
to accelerate the search over programs -- and (3) improving the DSL.
The key observation here is that each of these three steps can bootstrap off of each other:

@@

>>>>

Imagine an agent faced with a new domain of unfamiliar tasks: think
playing new kinds of videogames, authoring Python code, or navigating
through mazes.  Our agent has at its disposal a basis of primitive
actions it can compose to propose solutions to these tasks, but it is
no idea what kinds of primitives are appropriate for which tasks
nor does it know the higher-level vocabulary in which solutions are
best expressed.  How can our agent learn to solve problems in this new
domain?

The AI and machine learning literature contains two broad takes on this problem.
The first take is that the agent should come up with a better representation of the space of solutions,
for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
Our contribution is a general algorithm for fusing these two takes on the problem:
we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
alongside a bottom-up discriminative model that regresses from problems to solutions.
We evaluate our algorithm on four domains:
building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
Because we want a domain-general
approach,
we model solutions to tasks as
programs,
which frames our problem as one of program induction.

Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}: solid lines):
A DSL $\mathcal{D}$
probabilistically generates a latent program $p$
which generates the agent's observations, $x$.
In tandem we learn a learn a neural network $q$ that regresses
from observations to programs.

@@

@@

    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
  \end{scope}
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe tasks $x$, which are program inputs/outputs. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

Our goal in this work is to both induce a DSL and find good programs solving each of the tasks.
Our strategy is to iterate through three steps: (1) searching for programs that
solve the tasks, (2) learning a better neural recognition model -- which we use
to accelerate the search over programs -- and (3) improving the DSL.
The key observation here is that each of these three steps can bootstrap off each other:

@@

The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
learning a generative model and training a recognition model on samples from the generative model.
We apply this same idea to the problem of program induction,
and so we call our algorithm the \system.
